Optimal (Multiway) Spatial Joins

### 2024-06-01 msj.tex
yufei

\def\A{\mathcal{A}}
\def\B{\mathcal{B}}
\def\C{\mathcal{C}}
\def\E{\mathcal{E}}
\def\G{\mathcal{G}}
\def\I{\mathcal{I}}
\def\J{\mathcal{J}}
\def\II{\mathscr{I}}
\def\L{\mathcal{L}}
\def\P{\mathcal{P}}
\def\Q{\mathcal{Q}}
\def\R{\mathcal{R}}
\def\T{\mathcal{T}}
\def\U{\mathcal{U}}
\def\V{\mathcal{V}}
\def\X{\mathcal{X}}
\def\XX{\mathscr{X}}
\def\Y{\mathcal{Y}}
\def\YY{\mathscr{Y}}
\def\Z{\mathcal{Z}}

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\cat{\mathrm{cat1}}
\def\catt{\mathrm{cat2}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ... \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$. Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$. All logarithms have base 2 by default.

@@

\def\vgap{\vspace{2mm}}
\def\extraspacing{\vspace{2mm}}

@@

@@

Next, we consider the find-all-sorted version of Problem $\mathscr{D}$. Again, initialize an array $A$ of size $|R|$, where each rectangle of $R$ is allocated a cell. For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates. In the outset, all linked lists are empty. Unlike the detection version, we will not need to keep the active status for the rectangles.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} Sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$). Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position in the sorted list.

>>>>

\def\A{\mathcal{A}}
%\def\B{\mathcal{B}}
%\def\C{\mathcal{C}}
%\def\E{\mathcal{E}}
%\def\G{\mathcal{G}}
\def\I{\mathcal{I}}
\def\J{\mathcal{J}}
%\def\II{\mathscr{I}}
%\def\L{\mathcal{L}}
%\def\P{\mathcal{P}}
%\def\Q{\mathcal{Q}}
%\def\R{\mathcal{R}}
\def\T{\mathcal{T}}
%\def\U{\mathcal{U}}
%\def\V{\mathcal{V}}
%\def\X{\mathcal{X}}
%\def\XX{\mathscr{X}}
%\def\Y{\mathcal{Y}}
%\def\YY{\mathscr{Y}}
%\def\Z{\mathcal{Z}}

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\cat{\mathrm{cat1}}
\def\catt{\mathrm{cat2}}
\def\out{\mathrm{OUT}}
\def\dim{\mathrm{dim}}

@@

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ... \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$. Unless otherwise specified, every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$. All logarithms have base 2 by default.

@@

\def\vgap{\vspace{2mm}}
\def\extraspacing{\vspace{2mm} \noindent}

@@

Fix an integer $k \ge 3$, as well as $k$ integers $n_1, n_2, ..., n_k$ at least 1. Define $n = \sum_{i=1}^k n_i$. An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$.

A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a decision tree $\T$ defined as follows. Each internal node $u$ of $\T$ has three child nodes, referred as the {\em left}, {\em middle}, and {\em right} child, respectively. Each node is associated with an {\em instruction} of the form $(i_1, \dim_1, \lambda_1; i_2, \dim_2, \lambda_2)$.

@@

Next, we consider the find-all-sorted version of Problem $\mathscr{D}$. For each $r \in R$, we keep a linked list, which at the end of our algorithm will store the horizontal segments of $\cross_H(r)$ in ascending order of their y-coordinates. In the outset, all linked lists are empty. Unlike the detection version, we will not need to keep the active status for the rectangles.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$.} Sort all the rectangles $r \in R$ in ascending order of $\xright(r)$ (namely, the x-coordinate of the right edge of $r$). Assign the ID $i \in [R]$ to $r$ if $r$ is at the $i$-th position of the sorted list.
